---
title: 'ESM 262: Computing Assignment 5'
author: "Taylor Cook and Kelsey Warren"
date: "2025-03-11"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(devtools)
library(testthat)
library(purrr)
library(ggplot2)
library(dplyr)

```

# Design a function - you can pick any subject and you can even make up the equations as long as they conceptually make sense

```         
-   Make sure it has at least 2 inputs and 1 parameter (ideally
    more) and at least 2 outputs
-   Code your function in R, save as a R file in subdirectory called
    R
-   Make sure you include documentation (both at the top as we've
    shown in past examples, and inline)
-   Include some error checking
```

```{r}


source("~/Bren MESM/ESM 262 Computing/Assignment_5/R/oil.R")
oil

```

# In an Quatro file, generate some data for 2 of the function inputs

```         
-   use a *for* loop to run the function for the data
-   repeat the "looping" using something from the "purrr" package
-   Graph results (you can decide what the most interesting way to
    graph - you just need to make one graph)
    
    
```

```{r}

#Generate data for 2 of the function inputs
volume = c(100000, 200000, 300000, 400000, 500000)
wtemp = c(10, 15, 20, 25, 30)
windspeed <- rep(15, length(volume))

# run the function for the data using a for loop
results = list()
for (i in 1:length(volume)) {
  results[[i]] = oil(volume[i], wtemp[i], windspeed=15, viscosity=0.0001, ratecoeff=0.16)
}

print(results)

# repeat the "looping" using something from the "purrr" package
results_purrr = purrr::pmap(list(volume, wtemp, windspeed=15), ~oil(.x, .y, 15, viscosity=0.0001, ratecoeff=0.16))

print(results_purrr)

# Graph results of both methods

results_df = bind_rows(results) %>% 
  mutate(volume = volume, wtemp = wtemp, windspeed = windspeed)

results_purr_df = bind_rows(results_purrr) %>% 
  mutate(volume = volume, wtemp = wtemp, windspeed = windspeed)


results_df %>%
  ggplot(aes(x = volume, y = area_km2)) +
  geom_point() +
  geom_line() +
  labs(title = "Oil Spill Area vs. Volume", x = "Volume (m^3)", y = "Area (km^2)") +
  theme_minimal()


```

# Write at least 2 tests for your function; store in a separate test file

# Put this in a git repo that includes an R subdirectory, a tests subdirectory and your Quatro file and submit the link on Canvas (you can put all of this in a new git repo or a subdirectory of an existing git repo)
